Supporting Information for: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal

### 2020-09-17 Manuscript/SupportingInfo.tex

Here we examine the statistical properties of tests for comparing the
relative strength of phylogenetic signal, using \(\hat{Z}_{12}\).
Simulations were performed using pure-birth trees of different sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
combination of \(n\) and \(\lambda\), 101 traits were simulated on the
phylogeny under a Brownian motion model of evolution. Next, the
phylogenetic signal of trait 1 was compared to the signal in the
remaining 100 traits in pairwise fashion, using \(\hat{Z}_{12}\). The
proportion of significant results provided an estimate of the type I
error or model misspecification. Specifically, type I error was
evaluated when traits were simulated using no input phylogenetic signal
(i.e., \(\lambda = 0\)). Likewise, model misspecification was evaluated
for data simulated with some known, but equal, level of phylogenetic
signal (i.e., \(\lambda > 0\)). \hfill\break

\textbf{Results:} Tests revealed that across simulation conditions, the
average type I error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig.
S21). Levels of misspecification were slightly higher, and ranged on
average between 7\% (when comparing traits generated on the same
phylogeny: Fig. S21) to \textasciitilde12\% (for comparisons of traits
from different trees). Both results demonstrate that comparisons of the
strength of phylogenetic signal using \(\hat{Z}_{12}\) do not display
high levels of false positives, and are thus appropriate for comparing
the degree of phylogenetic signal across traits.

@@

\textbf{Figure S21}. Type I error rates and model misspecification for
\(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
(\(\lambda = 0\)).

>>>>

Here we examine the statistical properties of tests for comparing the
relative strength of phylogenetic signal, using \(\hat{Z}_{12}\).
Simulations were performed using pure-birth trees of different sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
combination of \(n\) and \(\lambda\), 50 traits were simulated on the
phylogeny under a Brownian motion model of evolution. Next, the
phylogenetic signal between traits was compared in pairwise fashion for
all \(50*49/2 = 1225\) combinations, using \(\hat{Z}_{12}\). The
proportion of significant results provided an estimate of the type I
error or model misspecification. Specifically, type I error was
evaluated when traits were simulated using no input phylogenetic signal
(i.e., \(\lambda = 0\)). Likewise, model misspecification was evaluated
for data simulated with some known, but equal, level of phylogenetic
signal (i.e., \(\lambda > 0\)). \hfill\break

\textbf{Results:} Tests revealed that across simulation conditions, the
average type I error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig.
S21). Levels of misspecification were slightly higher, and ranged on
average between 7\% when comparing traits generated on the same
phylogeny (Fig. S21 black line) to \textasciitilde12\% for comparisons
of traits from different trees. Both results demonstrate that
comparisons of the strength of phylogenetic signal using
\(\hat{Z}_{12}\) do not display high levels of false positives, and are
thus appropriate for comparing the degree of phylogenetic signal across
traits.

@@

\textbf{Figure S21}. Type I error rates and model misspecification for
\(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
(\(\lambda = 0\)). The solid black line represents the average type I
error and false discovery rate across all levels of input signal.
